### 2018-10-04 paper/casestudies.tex
further updates: use j instead of k in U_F(j)
Fixed errors found by Wen-ling

\begin{figure}[htbp]
\begin{center}
\begin{minipage}{0.3\textwidth} 
	 \begin{eqnarray*}
{ }[0]^0 & = & \{ a \} \\
{ }[1]^0 & = & \{ a,b,c \} \\
{ }[2]^0 & = & \{ a,b,c \} \\
{ }[3]^0 & = & \{ b,c \}
\end{eqnarray*}
	\end{minipage}
	\hfill
	\begin{minipage}{0.3\textwidth}
	 \begin{eqnarray*}
\minhits(0) & = & \{ \{ a\} \} \\
\minhits(1) & = & \{ \{a,b\}, \{c\}, \} \\
\minhits(2) & = & \{  \{a,b\},  \{a,c\},\} \\
\minhits(3) & = & \{ \{ b\}, \{c\} \} 
\end{eqnarray*}

	\end{minipage}
	\hfill
	\begin{minipage}{0.3\textwidth}
	 \begin{eqnarray*}
t(0,a) & = & 1 \\
t(1,a) & = & 0 \\
t(1,b) & = & 0 \\
t(1,c) & = & 2 \\
t(2,a) & = & 1 \\
t(2,b) & = & 0 \\
t(2,c) & = & 3 \\
t(3,b) & = & 0 \\
t(3,c) &  =& 3
\end{eqnarray*}
	\end{minipage}
\caption{Initials, minimal hitting sets, and transition function of the normalised transition graph displayed in Fig.~\ref{fig:tga}.}
\label{fig:initialsminhitstrans}
\end{center}
\end{figure}

@@

\begin{example}
\label{ex:uf1tests} Consider the following implementation $Z$ of process $P$
from Example~\ref{example:CSP} that is erroneous from the point of view of
failures refinement.
\begin{eqnarray*}
Z & = & a \then (Q_1 \intchoice R_1(r_{max},0))
\\
Q_1 & = & a\then Z \extchoice c\then Z
\\
R_1(r_{max},k) & = & (k < r_{max}) \& \big(b\then Z \extchoice  c\then R_1(r_{max},k+1)\big)
\\ & & \extchoice
\\ & & (k = r_{max}) \& \big(b\then Z \intchoice c\then R_1(r_{max},r_{max})\big)
\end{eqnarray*}
It is easy to see (and can be checked with FDR) that $Z$ is trace-equivalent
to $P$. While $k < r_{max}$, $Z$ also accepts the same sets of events as $P$.
When $R_1(r_{max},k)$ runs through several recursions and $k = r_{max}$,
however, $R_1(r_{max},k)$ makes an internal choice, instead of offering an
external choice, so $P\not\lessdet_F Z$. Fig.~\ref{fig:tgZ} shows the
normalised transition graph of $Z$ for $r_{max} = 3$.

Running the test $U_F(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p = 4$
states and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test depth
to be used according to Theorem~\ref{th:failurestest}), tests $U_F(0),\dots,
U_F(3)$ are passed by $Z$, but $Z$ fails $U_F(4)$, because after execution of
the trace
\[
s = a.c.c.c, \qquad\text{(note that $G(P)/s = \text{node}\ 3$ according to Fig.~\ref{fig:tga})},
\]
the test $U_F(4)$ offers hitting sets from $\minhits(3) = \{\{b\},\{c\}\}$
in branch (\ref{eq:ufd}). Therefore,  there exists one test execution
where $Z/s$ accepts only $\{b\}$ due to the internal choice (note
from Fig.~\ref{fig:tgZ} that
$G(Z)/s = 4$), while $U_F(4)/s$
only offers $\{c\}$ in branch (\ref{eq:ufd}) or $\{ a\} = \Sigma - [3]^0$ for
branch (\ref{eq:ufa}).  As a consequence, this execution of
$(Z\parallel[\Sigma] U_F(4))/s$ deadlocks, and the $pass$ event cannot be
produced. Another failing execution arises if $Z/s$ chooses to accept only
$\{c \}$, while $U_F(4)/s$ choses to accept only $\{a,b\}$. Therefore,
\fixme{The hiding is missing.} 
\fxnote{jp: done}
\[
(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F(4)) \hide\Sigma,
\]
and the test fails. \xbox
\end{example}

>>>>

\begin{figure}[htbp]
\begin{center}
\begin{minipage}{0.2\textwidth} 
	 \begin{eqnarray*}
{ }[0]^0 & = & \{ a \} \\
{ }[1]^0 & = & \{ a,b,c \} \\
{ }[2]^0 & = & \{ a,b,c \} \\
{ }[3]^0 & = & \{ b,c \}
\end{eqnarray*}
	\end{minipage}
	\hfill
	\begin{minipage}{0.33\textwidth}
	 \begin{eqnarray*}
\minhits(0) & = & \{ \{ a\} \} \\
\minhits(1) & = & \{ \{a,b\}, \{c\}, \} \\
\minhits(2) & = & \{  \{a,b\},  \{a,c\},\} \\
\minhits(3) & = & \{ \{ b\}, \{c\} \} 
\end{eqnarray*}

	\end{minipage}
	\hfill
	\begin{minipage}{0.2\textwidth}
	 \begin{eqnarray*}
t(0,a) & = & 1 \\
t(1,a) & = & 0 \\
t(1,b) & = & 0 \\
t(1,c) & = & 2  
\end{eqnarray*}
	\end{minipage}
		\hfill
	\begin{minipage}{0.2\textwidth}
	 \begin{eqnarray*}
t(2,a) & = & 1 \\
t(2,b) & = & 0 \\
t(2,c) & = & 3 \\
t(3,b) & = & 0 \\
t(3,c) &  =& 3
\end{eqnarray*}
	\end{minipage}

\caption{Initials, minimal hitting sets, and transition function of the normalised transition graph displayed in Fig.~\ref{fig:tga}.}
\label{fig:initialsminhitstrans}
\end{center}
\end{figure}

@@

\begin{example}
\label{ex:uf1tests} Consider the following implementation $Z$ of process $P$
from Example~\ref{example:CSP} that is erroneous from the point of view of
failures refinement. In the specification of $Z$, it is assumed that $r_{max}\ge 0$. 
\begin{eqnarray*}
Z & = & a \then (Q_1 \intchoice R_1(r_{max},0))
\\
Q_1 & = & a\then Z \extchoice c\then Z
\\
R_1(r_{max},k) & = & (k < r_{max}) \& \big(b\then Z \extchoice  c\then R_1(r_{max},k+1)\big)
\\ & & \extchoice
\\ & & (k = r_{max}) \& \big(b\then Z \intchoice c\then R_1(r_{max},r_{max})\big)
\end{eqnarray*}
It is easy to see (and can be checked with FDR) that $Z$ is trace-equivalent
to $P$. While $k < r_{max}$, $Z$ also accepts the same sets of events as $P$.
When $R_1(r_{max},k)$ runs through several recursions and $k = r_{max}$,
however, $R_1(r_{max},k)$ makes an internal choice, instead of offering an
external choice, so $P\not\lessdet_F Z$. Fig.~\ref{fig:tgZ} shows the
normalised transition graph of $Z$ for $r_{max} = 3$.

Running the test $U_F(j)$ against $Z$ for $j=0,\dots,19$ ($G(P)$ has $p = 4$
states and $G(Z)$ has $q=5$, so $pq-1=19$ is the index of the last test to 
be executed  according to Theorem~\ref{th:failurestest}), tests $U_F(0),\dots,
U_F(3)$ are passed by $Z$, but $Z$ fails $U_F(4)$, because after execution of
the trace
\[
s = a.c.c.c, \qquad\text{(note that $G(P)/s = \text{node}\ 3$ according to Fig.~\ref{fig:tga})},
\]
the test $U_F(4)$ offers hitting sets from $\minhits(3) = \{\{b\},\{c\}\}$
in branch (\ref{eq:ufd}). Therefore,  there exists one test execution
where $Z/s$ accepts only $\{b\}$ due to the internal choice (note
from Fig.~\ref{fig:tgZ} that
$G(Z)/s = \text{node}\ 4$), while $U_F(4)/s$
only offers $\{c\}$ in branch (\ref{eq:ufd}) or $\{ a\} = \Sigma - [3]^0$ for
branch (\ref{eq:ufa}).  As a consequence, this execution of
$(Z\parallel[\Sigma] U_F(4))/s$ deadlocks, and the $pass$ event cannot be
produced. Another failing execution arises if $Z/s$ chooses to accept only
$\{c \}$, while $U_F(4)/s$ choses to accept only $\{a,b\}$. Therefore,
\fixme{The hiding is missing.} 
\fxnote{jp: done}
\[
(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F(4)) \hide\Sigma,
\]
and the test fails. \xbox
\end{example}

### 2018-10-04 paper/complexity.tex
Fixed errors found by Wen-ling

Process $Q$, however, neither conforms to $P$ in the failures refinement 
relation, nor in the trace
refinement relation. This can only be seen when executing the test $U_F(pq-1)$ and
$U_T(pq-1)$, respectively. These test fail, so this shows that $P\not\lessdet_F Q$ and
$P\not\lessdet_F Q$ according to Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}. Moreover, this shows that
the maximal trace
length $pq$ to be investigated in the tests cannot be further reduced without losing
the completeness property of the test suites.
\xbox
\end{example}

@@

To prove statement~5, we observe that the specification of $P$ implies
\[
\minaccs(P/s) = \left\{
\begin{array}{ll}
\{ \{a\}, \{b\} \} & \text{for all $s\in\trc(P)$ with $\#s<p-1$.}
\\
\{ \{a\} \} & \text{for all $s\in\trc(P)$ with $\#s\ge p-1$.}
\end{array}
\right.
\]
and
\[
\minaccs(Q/s) = \left\{
\begin{array}{ll}
\{ \{a\}  \} & \text{for all $s\in\trc(Q)$ with $\#s \neq 0\mod(q-1)$.}
\\
\{ \{a,b\} \} & \text{for all $s\in\trc(P)$ with $\#s = 0\mod(q-1)$.}
\end{array}
\right.
\]
As a  consequence, the minimal acceptance set $A_P = \{a\}$ which is
contained in every $\minaccs(P/s)$ fulfils $A_P \subseteq A_Q$ for
any $A_Q\in\minaccs(Q/s)$, when $s\in\trc(P)\cap \trc(Q)$. Now Lemma~\ref{lemma:tgtrcref}, 
(\ref{eq:failrefb}) can be applied to conclude that $Q\ conf\ P$.
\xbox
\end{proof}
%
Since Theorem~\ref{th:maxtracelen} just states that a violation of trace
refinement may remain undetected if only traces shorter than $pq$ are checked
during tests, it can also be applied to our trace refinement tests.
Therefore, test suites $\{ U_T(z) \}$ with $z<pq-1$ are not complete. It is
discussed in Section~\ref{sec:conc} how the number of test traces to be
executed by complete test suites for failures or trace refinement can still
be reduced {\it without} reducing the maximal length.

>>>>

Process $Q$, however, neither conforms to $P$ in the failures refinement 
relation, nor in the trace
refinement relation. This can only be seen when executing the test $U_F(pq-1)$ and
$U_T(pq-1)$, respectively. These tests fail, so this shows that $P\not\lessdet_F Q$ and
$P\not\lessdet_T Q$ according to Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}. Moreover, this shows that
the maximal trace
length $pq$ to be investigated in the tests cannot be further reduced without losing
the completeness property of the test suites.
\xbox
\end{example}

@@

To prove statement~5, we observe that the specification of $P$ implies (the 
expression $(s\cnt b)$ denotes the number of $b$-events occurring in trace $s$)
\[
\minaccs(P/s) = \left\{
\begin{array}{ll}
\{ \{a\}, \{b\} \} & \text{for all $s\in\trc(P)$ with $(s\cnt b) <p-1$.}
\\
\{ \{a\} \} & \text{for all $s\in\trc(P)$ with $(s\cnt b) = p-1$.}
\end{array}
\right.
\]
and
\[
\minaccs(Q/s) = \left\{
\begin{array}{ll}
\{ \{a\}  \} & \text{for all $s\in\trc(Q)$ with $\#s \neq 0\mod(q-1)$.}
\\
\{ \{a,b\} \} & \text{for all $s\in\trc(P)$ with $\#s = 0\mod(q-1)$.}
\end{array}
\right.
\]
As a  consequence, the minimal acceptance set $A_P = \{a\}$ which is
contained in every $\minaccs(P/s)$ fulfils $A_P \subseteq A_Q$ for
any $A_Q\in\minaccs(Q/s)$, when $s\in\trc(P)\cap \trc(Q)$. Now Lemma~\ref{lemma:tgtrcref}, 
(\ref{eq:failrefb}) can be applied to conclude that $Q\ conf\ P$.
\xbox
\end{proof}
%
Since Theorem~\ref{th:maxtracelen} just states that a violation of trace
refinement may remain undetected if only traces shorter than $pq$ are checked
during tests, it can also be applied to our trace refinement tests.
Therefore, test suites $\{ U_T(z) \}$ with $z<pq-1$ are not complete. It is
discussed in Section~\ref{sec:conc} how the number of test traces to be
executed by complete test suites for failures or trace refinement can still
be reduced {\it without} reducing the maximal length.
